---
title: "Analysing and visualising rating scales in R using ratingScaleSummary"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

ratingScaleSummary provides tools for quick analysis and visualisation of rating scales,
consisting of several variables using the same categories/
codings. These procedures might be useful for social scientists and psychologists
and everyone else working with data collected via surveys making use of the the
concept of multiple indicators for the measurment of dimensions of abstract or
complex social concepts and terms. 

In addition there are functions for calculating the interpolated median
and percentiles of a grouped frequency distribution, the inter quartile range,
and relative frequency and cross tables. The latter functions produce data
frames, the output is therefore modified and reused much easier, than the
outputs created by, e.g. the functions of the package 'descr'.

The functions in this package produce output that is supposed to be easily integrated into
RMarkdown and knitr files. All tabular outputs are **data frames** which integrate nicely into your
workflow.

The plots created by this package could be called 'vertical-line-(dot) plots' (If there is another
or a better name, please let me know!). Vertical-line plots can be understood intuitivly, since
they resemble the structure of a typical rating scale layout, as it is presented to the survey 
participants. When used for group comparison vertical-line plots give a quick insight into the 
variables that 'stir controversy' between the groups, or which don't.

# Install Package

For now the ratingScaleSummary package is not available on CRAN, but it can be
installed directly from github within R using the devtools package:

    devtools::install_github('ratingScaleSummary/tilltnet')

or by using the source package for a local install.

# Let's get started: creating sample data

We start with creating some sample data, that will be used throughout this
vignette. We'll need a *data.frame* consisting of several variables, all using the
same categories. In addition we create a grouping variable, that is used for
the examples that incorporate group comparisson.

```{r}
df <- data.frame(replicate(6,sample(as.character(1:7), 2000, replace = T, prob = sample(1:100/100, 7))))
var_names <- lapply(1:6, FUN = function(x) paste(sample(letters, sample(6:13, 1)), collapse = ""))
names(df) <- var_names

g <- sample(c("G1", "G2", "G3"), 2000, replace = T)
```


# Summarise a rating scale

rs_summary() returns a **data.frame** where each row
represents one variable. It provides categories' frequencies (in %),
N, median, and the interquartile range.
```{r}
library(ratingScaleSummary)
res <- rs_summary(df = df)
knitr::kable(res)
```

The rs_summary() command gives a summary of the

## Visualise medians of a rating scale

vert_line_base() and vert_line_gg() plots a vertical line dot plot using base plotting or ggplot2.

```{r, fig.height=5, fig.width=7}
# Using base plot.
vert_line_base(x = res["Median"], max_val = 7)
# Using ggplot2.
vert_line_gg(x = res["Median"], max_val = 7)
```

# Compare the responses to a rating scale by groups

Returns a **data.frame** where each row represents one variable
and the columns show the median values of the comparison groups. An additional
colum shows the significance levels, based on the **Kruskal-Wallis** **Rank**
**Sum** **Test**.


```{r}
# Compare medians between groups accross several variables.
res_comp <- rs_compare_groups(df = df, groups = g)
knitr::kable(res_comp)
```

## Visualise the comparisson

vert_line_base() and vert_line_gg() also produce plots for group comparisson.

```{r, fig.height=5, fig.width=7}
# Using base plot.
vert_line_base(x = res_comp, max_val = 7)
# Using ggplot2.
vert_line_gg(x = res_comp, max_val = 7)
```

# Using ggplot2 theming

```{r, fig.height=5, fig.width=7}
library(ggthemes)
vert_line_gg(x = res["Median"], max_val = 7) +  theme_hc()
vert_line_gg(x = res_comp, max_val = 7) + theme_wsj()
```


# Miscellaneous functions within the package

 - gmedian() calculates the interpolated median of a grouped
   frequency distribution.
 - intQR() calculates interquartile range of a grouped frequency
   distribution.
 - mQ() Calculates halfed interquartile range [german: mittlerer 
   Quartilsabstand] of a grouped frequency distribution
 - f_tab() returns a **data frame** where each row shows the
   frequencies of unique values in a vector. Weighting is supported.
 - x_tab() returns a **data frame** as a crosstable with indication of 
   statistic significance using chi-square. Weighting is supported.
